---
title: "TSA_FinalProject_Coding"
output:
  pdf_document: default
  html_document: default
date: "2023-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## install packages

```{r package, message=FALSE, include=FALSE}
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(kableExtra)
library(knitr)
library(forecast)

```

## Dataset Information

```{r}
#Import the excel file
library(readxl)
electricity_consumption <- read_excel(path="../Table_2.1a__Energy_Consumption__Residential_-_Commercial_-_and_Industrial_Sectors.xlsx",skip = 10, sheet="Monthly Data",col_names=TRUE) 

#delete the first row with units
electricity_consumption<-electricity_consumption[-1,]

#Keep two columns: primary energy consumption and energy losses for three sectors, respectively
electricity_consumption<-electricity_consumption[,-3][,-3][,-4]
electricity_consumption<-electricity_consumption[,-5][,-5][,-6]
electricity_consumption<-electricity_consumption[,-7][,-7][,-8]

#Round to three digits for all columns
class(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)
electricity_consumption$`Primary Energy Consumed by the Residential Sector` <- as.numeric(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)
round(electricity_consumption$`Primary Energy Consumed by the Residential Sector`,digits=3)

class(electricity_consumption$`Residential Sector Electrical System Energy Losses`)
electricity_consumption$`Residential Sector Electrical System Energy Losses` <- as.numeric(electricity_consumption$`Residential Sector Electrical System Energy Losses`)
round(electricity_consumption$`Residential Sector Electrical System Energy Losses`,digits=3)

class(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)
electricity_consumption$`Primary Energy Consumed by the Commercial Sector` <- as.numeric(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)
round(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`,digits=3)

class(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)
electricity_consumption$`Commercial Sector Electrical System Energy Losses` <- as.numeric(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)
round(electricity_consumption$`Commercial Sector Electrical System Energy Losses`,digits=3)

class(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)
electricity_consumption$`Primary Energy Consumed by the Industrial Sector` <- as.numeric(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)
round(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`,digits=3)

class(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
electricity_consumption$`Industrial Sector Electrical System Energy Losses` <- as.numeric(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
round(electricity_consumption$`Industrial Sector Electrical System Energy Losses`,digits=3)

colnames(electricity_consumption)<-c('Date','Residential.Primary.Energy','Residential.Energy.Losses','Commercial.Primary.Energy','Commercial.Energy.Losses','Industrial.Primary.Energy','Industrial.Energy.Losses')

head(electricity_consumption)
```

```{r message=FALSE, warning=FALSE}
#Transform to time series data
Date_frame<-as.Date(electricity_consumption$Month)
ts_electricity <- ts(electricity_consumption[,2:7],start=c(1973,01),end=c(2022,11),frequency=12)

#Summary of time series data
summary_data<-summary(ts_electricity)
summary_data

#Calculate means, standard deviations, and ranges for each variable
mean_R1<-mean(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)
sd_R1<-sd(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)
max_R1<-max(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)
min_R1<-min(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)
range_R1<-max(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)-min(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)

mean_R2<-mean(electricity_consumption$`Residential Sector Electrical System Energy Losses`)
sd_R2<-sd(electricity_consumption$`Residential Sector Electrical System Energy Losses`)
max_R2<-max(electricity_consumption$`Residential Sector Electrical System Energy Losses`)
min_R2<-min(electricity_consumption$`Residential Sector Electrical System Energy Losses`)
range_R2<-max(electricity_consumption$`Residential Sector Electrical System Energy Losses`)-min(electricity_consumption$`Residential Sector Electrical System Energy Losses`)

mean_C1<-mean(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)
sd_C1<-sd(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)
max_C1<- max(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)
min_C1<- min(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)
range_C1<-max(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)-min(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)

mean_C2<-mean(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)
sd_C2<-sd(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)
max_C2<- max(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)
min_C2<- min(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)
range_C2<-max(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)-min(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)

mean_I1<-mean(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)
sd_I1<-sd(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)
max_I1<-max(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)
min_I1<-min(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)
range_I1<-max(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)-min(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)

mean_I2<-mean(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
sd_I2<-sd(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
max_I2<-max(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
min_I2<-min(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
range_I2<-max(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)-min(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
```


```{r}
# Create a table
summary_df <- data.frame(
  Variable = c("Residential_Sector_Consumption", "Residential_Sector_Losses", "Commercial_Sector_Consumption", "Commercial_Sector_Losses", "Industrial_Sector_Consumption", "Industrial_Sector_Losses"),
  Mean = c(584.60,662.45,345.14,605.47,1806.54,576.51),
  StandardDeviation = c(321.54,185.03,146.90,178.14,146.71,68.86),
  Max = c(1488.0, 1144.8,703.6,982.2,2267,776.8),
  Min = c(192.3,325.3,165.3,258.2,1466,405.0),
  Range = c(1295.71,819.46,538.31,723.97,800.93,371.76)
)
  
summary_table<-kbl(summary_df, caption="Summary Statistics Table used with ts() function") %>% 
  kable_styling(full_width = FALSE, position = "center",latex_options = "hold_position")
summary_table
```


## Initial Plots

```{r}
#plot for electricity consumption
cols<-c("residential"="orange","commercial"="steelblue","industrial"="yellow")
ts_plot<-ggplot(electricity_consumption)+
  geom_line(aes(x=Date,y=Residential.Primary.Energy,color="residential"))+
  ylab(paste0("Electricity Consumption (Trillion BTU)",sep="")) +
  xlab("Date")+
  geom_line(aes(x=Date,y=Commercial.Primary.Energy,color="commercial"))+
  geom_line(aes(x=Date,y=Industrial.Primary.Energy,color="industrial"))+
  labs(title="Primary Electricity Consumption of Three Sectors")+
  scale_color_manual(name="series",values = cols)
plot(ts_plot)
          
#plot for energy losses
cols1<-c("residential"="green","commercial"="red","industrial"="blue")
ts_plot1<-ggplot(electricity_consumption)+
  geom_line(aes(x=Date,y=Residential.Energy.Losses,color="residential"))+
  ylab(paste0("Energy Losses (Trillion BTU)",sep="")) +
  xlab("Date")+
  geom_line(aes(x=Date,y=Commercial.Energy.Losses,color="commercial"))+
  geom_line(aes(x=Date,y=Industrial.Energy.Losses,color="industrial"))+
  labs(title="Energy Losses of Three Sectors")+
  scale_color_manual(name="series",values = cols1)
plot(ts_plot1)

```

###Residential Sector
```{r}
#Create a dataframe only for Residential
electricity_consumption_R<-electricity_consumption[,-4:-7]
colnames(electricity_consumption_R)<-c('Date','Primary.Energy','Energy.Losses')

#Inspect data
#head(electricity_price)
nvar <- ncol(electricity_consumption_R) - 1
nobs <- nrow(electricity_consumption_R)

#transform to time series data
ts_electricity_R <- ts(
  electricity_consumption_R[,2:(nvar+1)],
  start=c(1973,11,01),end=c(2022,11,01),
  frequency=12)

```

```{r}
# Create a variable with number of steps ahead we will forecast
n_for <- 12

ts_forPE_R_train <- ts(ts_electricity_R[1:(nobs-n_for),"Primary.Energy"],
               start=c(1973,11,01),end=c(2021,11,01),
               frequency=12)

last_forPE_obs_R_test <- ts(ts_electricity_R[(nobs-n_for):589,"Primary.Energy"],
               start=c(2021,12,01),end=c(2022,11,01),
               frequency=12)

ts_forEL_R_train <- ts(ts_electricity_R[1:(nobs-n_for),"Energy.Losses"],
               start=c(1973,11,01),end=c(2021,11,01),
               frequency=12)
last_EL_obs_R_test <- ts(ts_electricity_R[(nobs-n_for):589,"Energy.Losses"],
               start=c(2021,12,01),end=c(2022,11,01),
               frequency=12)

```


##ACF and PACF Plots
```{r}
#ACF and PACF plots
par(mfrow=c(1,4))
ACF_Plot <- Acf(electricity_consumption_R$Primary.Energy, lag = 40, plot = TRUE,main="ACF of Residential Sector Consumption")
PACF_Plot <- Pacf(electricity_consumption_R$Primary.Energy, lag = 40, plot = TRUE,main="PACF of Residential Sector Consumption")
ACF_Plot <- Acf(electricity_consumption_R$Energy.Losses, lag = 40, plot = TRUE,main="ACF of Residential Sector Losses")
PACF_Plot <- Pacf(electricity_consumption_R$Energy.Losses, lag = 40, plot = TRUE,main="PACF of Residential Sector Losses")
```

## Model 1: Arithmetic mean
```{r}
#mean model for primary energy consumption
MEAN_seas1 <- meanf(y = ts_forPE_R_train, h = n_for)  
checkresiduals(MEAN_seas1)
plot(MEAN_seas1)


#mean model for energy losses
MEAN_seas2 <- meanf(y = ts_forEL_R_train, h = n_for)  
checkresiduals(MEAN_seas2)
plot(MEAN_seas2)
```

## Model 2: Seasonal naive
```{r}
#seasonal naive model for primary energy consumption
SNAIVE_seas1 <- snaive(ts_forPE_R_train, h=n_for, holdout=FALSE)
checkresiduals(SNAIVE_seas1)
plot(SNAIVE_seas1)

#seasonal naive model for energy losses
SNAIVE_seas2 <- snaive(ts_forEL_R_train, h=n_for, holdout=FALSE)
checkresiduals(SNAIVE_seas2)
plot(SNAIVE_seas2)
```

## Model 3: SARIMA
```{r}
##SARIMA model for primary energy consumption
SARIMA_autofit1 <- auto.arima(ts_forPE_R_train)
checkresiduals(SARIMA_autofit1)

SARIMA_for1 <- forecast(SARIMA_autofit1,h=n_for)
plot(SARIMA_for1)

##SARIMA model for energy losses
SARIMA_autofit2 <- auto.arima(ts_forEL_R_train)
checkresiduals(SARIMA_autofit2)

SARIMA_for2 <- forecast(SARIMA_autofit2,h=n_for)
plot(SARIMA_for2)
```

## Fitting State Space Models to the original (seasonal) series
## Model 4: SS Exponential smoothing

```{r}
##SSEAS model for primary energy consumption
SSES_seas1 <- es(ts_forPE_R_train,model="ZZZ",h=n_for,holdout=FALSE)
plot(SSES_seas1)
checkresiduals(SSES_seas1)

##SSEAS model for energy losses
SSES_seas2 <- es(ts_forEL_R_train,model="XXX",h=n_for,holdout=FALSE)
plot(SSES_seas2)
checkresiduals(SSES_seas2)
```

### Model 5: SS with StructTS()


```{r}
#For primary energy consumption
SS_seas1 <- StructTS(ts_forPE_R_train,
                    type="BSM",fixed=c(NA,NA,NA,NA))
checkresiduals(SS_seas1)

#For energy losses
SS_seas2 <- StructTS(ts_forEL_R_train,
                    type="BSM",fixed=c(NA,NA,NA,NA))
checkresiduals(SS_seas2)

#Generating forecasts
SS_for1 <- forecast(SS_seas1,h=n_for)
plot(SS_for1)

SS_for2 <- forecast(SS_seas2,h=n_for)
plot(SS_for2)
```


## Checking accuracy of the five models

```{r}
#Model 1: Arithmetic mean
MEAN_scores1 <- accuracy(MEAN_seas1$mean,last_forPE_obs_R_test)
MEAN_scores2 <- accuracy(MEAN_seas2$mean,last_EL_obs_R_test)

#Model 2: Seasonal naive 
SNAIVE_scores1 <- accuracy(SNAIVE_seas1$mean,last_forPE_obs_R_test)
SNAIVE_scores2 <- accuracy(SNAIVE_seas2$mean,last_EL_obs_R_test)

# Model 3:  SARIMA 
SARIMA_scores1 <- accuracy(SARIMA_for1$mean,last_forPE_obs_R_test)
SARIMA_scores2 <- accuracy(SARIMA_for2$mean,last_EL_obs_R_test)

# Model 4:  SSES
SSES_scores1 <- accuracy(SSES_seas1$forecast,last_forPE_obs_R_test)
SSES_scores2 <- accuracy(SSES_seas2$forecast,last_EL_obs_R_test)

# Model 5:  BSM 
SS_scores1 <- accuracy(SS_for1$mean,last_forPE_obs_R_test)
SS_scores2 <- accuracy(SS_for2$mean,last_EL_obs_R_test)
```

## Compare performance metrics

````{r}
#create data frame for primary energy consumption
seas_scores1 <- as.data.frame(rbind(MEAN_scores1, SNAIVE_scores1, SARIMA_scores1,SSES_scores1,SS_scores1))
row.names(seas_scores1) <- c("MEAN", "SNAIVE","SARIMA","SSES","BSM")

#choose model with lowest RMSE
best_model_index1 <- which.min(seas_scores1[,"RMSE"])
cat("The best model by RMSE is:", row.names(seas_scores1[best_model_index1,]))                       
                            
```

```{r}

kbl_PE_R<-kbl(seas_scores1, 
      caption = "Forecast Accuracy for Residential Primary Energy Consumption's Seasonal Data",
      digits = array(5,ncol(seas_scores1))) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  kable_styling(latex_options="striped", stripe_index = which.min(seas_scores1[,"RMSE"]))

kbl_PE_R
```

````{r}
#create data frame for residential energy losses
seas_scores2 <- as.data.frame(rbind(MEAN_scores2, SNAIVE_scores2, SARIMA_scores2,SSES_scores2,SS_scores2))
row.names(seas_scores2) <- c("MEAN", "SNAIVE","SARIMA","SSES","BSM")

#choose model with lowest RMSE
best_model_index2 <- which.min(seas_scores2[,"RMSE"])
cat("The best model by RMSE is:", row.names(seas_scores2[best_model_index2,]))                       
                            
```

```{r}

kbl_EL_R<-kbl(seas_scores2, 
      caption = "Forecast Accuracy for Residential Enegy Losses' Seasonal Data",
      digits = array(5,ncol(seas_scores2))) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  kable_styling(latex_options="striped", stripe_index = which.min(seas_scores2[,"RMSE"]))

kbl_EL_R
```

## Plotting everything together

```{r}
#Residential Primary Energy
autoplot(ts_electricity_R[,"Primary.Energy"]) +
  autolayer(MEAN_seas1, PI=FALSE, series="Mean") +
  autolayer(SNAIVE_seas1, PI=FALSE, series="Na誰ve") +
  autolayer(SARIMA_for1,PI=FALSE, series="SARIMA") +
  autolayer(SSES_seas1$forecast, series="SSES") +
  autolayer(SS_for1,PI=FALSE,series="BSM") + 
  xlab("Month") + ylab("Electricity Consumption (Trillion BTU)") +
  guides(colour=guide_legend(title="Forecast"))
```

```{r}
#Residential Energy Losses
autoplot(ts_electricity_R[,"Energy.Losses"]) +
  autolayer(MEAN_seas2, PI=FALSE, series="Mean") +
  autolayer(SNAIVE_seas2, PI=FALSE, series="Na誰ve") +
  autolayer(SARIMA_for2,PI=FALSE, series="SARIMA") +
  autolayer(SSES_seas2$forecast, series="SSES") +
  autolayer(SS_for2,PI=FALSE,series="BSM") + 
  xlab("Month") + ylab("Electricity Losses (Trillion BTU)") +
  guides(colour=guide_legend(title="Forecast"))
```
###Commercial Sector
```{r}

#Create a dataframe only for Residential
electricity_consumption_C<-electricity_consumption[c(1,4,5)]
colnames(electricity_consumption_C)<-c('Date','Primary.Energy','Energy.Losses')

#Inspect data
#head(electricity_price)
nvar <- ncol(electricity_consumption_C) - 1
nobs <- nrow(electricity_consumption_C)

#transform to time series data
ts_electricity_C <- ts(
  electricity_consumption_C[,2:(nvar+1)],
  start=c(1973,11,01),end=c(2022,11,01),
  frequency=12)

```


```{r}
# Create a variable with number of steps ahead we will forecast
n_for <- 12

ts_forPE_C_train <- ts(ts_electricity_C[1:(nobs-n_for),"Primary.Energy"],
               start=c(1973,11,01),end=c(2021,11,01),
               frequency=12)

last_forPE_obs_C_test <- ts(ts_electricity_C[(nobs-n_for):589,"Primary.Energy"],
               start=c(2021,12,01),end=c(2022,11,01),
               frequency=12)

ts_forEL_C_train <- ts(ts_electricity_C[1:(nobs-n_for),"Energy.Losses"],
               start=c(1973,11,01),end=c(2021,11,01),
               frequency=12)
last_EL_obs_C_test <- ts(ts_electricity_C[(nobs-n_for):589,"Energy.Losses"],
               start=c(2021,12,01),end=c(2022,11,01),
               frequency=12)
```


##ACF and PACF Plots
```{r}
#ACF and PACF plots
par(mfrow=c(1,4))
ACF_Plot <- Acf(electricity_consumption_C$Primary.Energy, lag = 40, plot = TRUE,main="ACF of Commercial Sector Consumption")
PACF_Plot <- Pacf(electricity_consumption_C$Primary.Energy, lag = 40, plot = TRUE,main="PACF of Commercial Sector Consumption")
ACF_Plot <- Acf(electricity_consumption_C$Energy.Losses, lag = 40, plot = TRUE,main="ACF of Commercial Sector Losses")
PACF_Plot <- Pacf(electricity_consumption_C$Energy.Losses, lag = 40, plot = TRUE,main="PACF of Commercial Sector Losses")
```

## Model 1: Arithmetic mean
```{r}
#mean model for primary energy consumption
MEAN_seas1_C <- meanf(y = ts_forPE_C_train, h = n_for)  
checkresiduals(MEAN_seas1_C)
plot(MEAN_seas1_C)

#mean model for energy losses
MEAN_seas2_C <- meanf(y = ts_forEL_C_train, h = n_for)  
checkresiduals(MEAN_seas2_C)
plot(MEAN_seas2_C)
```

## Model 2: Seasonal naive
```{r}
#seasonal naive model for primary energy consumption
SNAIVE_seas1_C <- snaive(ts_forPE_C_train, h=n_for, holdout=FALSE)
checkresiduals(SNAIVE_seas1_C)
plot(SNAIVE_seas1_C)

#seasonal naive model for energy losses
SNAIVE_seas2_C <- snaive(ts_forEL_C_train, h=n_for, holdout=FALSE)
checkresiduals(SNAIVE_seas2_C)
plot(SNAIVE_seas2_C)
```

## Model 3: SARIMA
```{r}
##SARIMA model for primary energy consumption
SARIMA_autofit1_C <- auto.arima(ts_forPE_C_train)
checkresiduals(SARIMA_autofit1_C)

SARIMA_for1_C <- forecast(SARIMA_autofit1_C,h=n_for)
plot(SARIMA_for1_C)

##SARIMA model for energy losses
SARIMA_autofit2_C <- auto.arima(ts_forEL_C_train)
checkresiduals(SARIMA_autofit2_C)

SARIMA_for2_C <- forecast(SARIMA_autofit2_C,h=n_for)
plot(SARIMA_for2_C)
```

## Fitting State Space Models to the original (seasonal) series
## Model 4: SS Exponential smoothing

```{r}
##SSEAS model for primary energy consumption
SSES_seas1_C <- es(ts_forPE_C_train,model="ZZZ",h=n_for,holdout=FALSE)
plot(SSES_seas1_C)
checkresiduals(SSES_seas1_C)

##SSEAS model for energy losses
SSES_seas2_C <- es(ts_forEL_C_train,model="XXX",h=n_for,holdout=FALSE)
plot(SSES_seas2_C)
checkresiduals(SSES_seas2_C)
```

### Model 5: SS with StructTS()


```{r}
#For primary energy consumption
SS_seas1_C <- StructTS(ts_forPE_C_train,
                    type="BSM",fixed=c(NA,NA,NA,NA))
checkresiduals(SS_seas1_C)

#For energy losses
 SS_seas2_C<- StructTS(ts_forEL_C_train,
                    type="BSM",fixed=c(NA,NA,NA,NA))
checkresiduals(SS_seas2_C)

#Generating forecasts
SS_for1_C <- forecast(SS_seas1_C,h=n_for)
plot(SS_for1_C)

SS_for2_C <- forecast(SS_seas2_C,h=n_for)
plot(SS_for2_C)
```


## Checking accuracy of the five models

```{r}
#Model 1: Arithmetic mean
MEAN_scores1_C <- accuracy(MEAN_seas1_C$mean,last_forPE_obs_C_test)
MEAN_scores2_C <- accuracy(MEAN_seas2_C$mean,last_EL_obs_C_test)

#Model 2: Seasonal naive 
SNAIVE_scores1_C<- accuracy(SNAIVE_seas1_C$mean,last_forPE_obs_C_test)
SNAIVE_scores2_C <- accuracy(SNAIVE_seas2_C$mean,last_EL_obs_C_test)

# Model 3:  SARIMA 
SARIMA_scores1_C <- accuracy(SARIMA_for1_C$mean,last_forPE_obs_C_test)
SARIMA_scores2_C <- accuracy(SARIMA_for2_C$mean,last_EL_obs_C_test)

# Model 4:  SSES
SSES_scores1_C <- accuracy(SSES_seas1_C$forecast,last_forPE_obs_C_test)
SSES_scores2_C <- accuracy(SSES_seas2_C$forecast,last_EL_obs_C_test)

# Model 5:  BSM 
SS_scores1_C <- accuracy(SS_for1_C$mean,last_forPE_obs_C_test)
SS_scores2_C <- accuracy(SS_for2_C$mean,last_EL_obs_C_test)
```

```{r}
#create data frame for primary energy consumption
seas_scores1_C <- as.data.frame(rbind(MEAN_scores1_C, SNAIVE_scores1_C, SARIMA_scores1_C,SSES_scores1,SS_scores1_C))
row.names(seas_scores1_C) <- c("MEAN", "SNAIVE","SARIMA","SSES","BSM")
#choose model with lowest RMSE
best_model_index1_C <- which.min(seas_scores1_C[,"RMSE"])
cat("The best model by RMSE is:", row.names(seas_scores1_C[best_model_index1_C,]))                       
                            
```

```{r}
kbl_PE_C<-kbl(seas_scores1_C, 
      caption = "Forecast Accuracy for Commercial Primary Energy Consumption's Seasonal Data",
      digits = array(5,ncol(seas_scores1_C))) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  kable_styling(latex_options="striped", stripe_index = which.min(seas_scores1_C[,"RMSE"]))
kbl_PE_C
```

```{r}
seas_scores2_C <- as.data.frame(rbind(MEAN_scores2_C, SNAIVE_scores2_C, SARIMA_scores2_C,SSES_scores2_C,SS_scores2_C))
row.names(seas_scores2_C) <- c("MEAN", "SNAIVE","SARIMA","SSES","BSM")
#choose model with lowest RMSE
best_model_index2_C <- which.min(seas_scores2_C[,"RMSE"])
cat("The best model by RMSE is:", row.names(seas_scores2_C[best_model_index2_C,])) 
```

```{r}
kbl_EL_C<-kbl(seas_scores2_C, 
      caption = "Forecast Accuracy for Commercial Enegy Losses' Seasonal Data",
      digits = array(5,ncol(seas_scores2_C))) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  kable_styling(latex_options="striped", stripe_index = which.min(seas_scores2_C[,"RMSE"]))
kbl_EL_C
```

```{r}
## Plotting everything together
#Commercial Primary Energy
autoplot(ts_electricity_C[,"Primary.Energy"]) +
  autolayer(MEAN_seas1_C, PI=FALSE, series="Mean") +
  autolayer(SNAIVE_seas1_C, PI=FALSE, series="Na誰ve") +
  autolayer(SARIMA_for1_C,PI=FALSE, series="SARIMA") +
  autolayer(SSES_seas1_C$forecast, series="SSES") +
  autolayer(SS_for1_C,PI=FALSE,series="BSM") + 
  xlab("Month") + ylab("Electricity Consumption (Trillion BTU)") +
  guides(colour=guide_legend(title="Forecast"))
```

```{r}
#Commercial Energy Losses
autoplot(ts_electricity_C[,"Energy.Losses"]) +
  autolayer(MEAN_seas2_C, PI=FALSE, series="Mean") +
  autolayer(SNAIVE_seas2_C, PI=FALSE, series="Na誰ve") +
  autolayer(SARIMA_for2_C,PI=FALSE, series="SARIMA") +
  autolayer(SSES_seas2_C$forecast, series="SSES") +
  autolayer(SS_for2_C,PI=FALSE,series="BSM") + 
  xlab("Month") + ylab("Electricity Losses (Trillion BTU)") +
  guides(colour=guide_legend(title="Forecast"))
```


####Industrial Sector

```{r}
I_electricity_consumption<-electricity_consumption[c(1,6,7)]
colnames(electricity_consumption_R)<-c('Date','Industrial.Primary.Energy','Energy.Losses')
#Inspect data
#head(electricity_price)
nvar <- ncol(I_electricity_consumption) - 1
nobs <- nrow(I_electricity_consumption)
#transform 
I_ts_electricity <- ts(
  I_electricity_consumption[,2:(nvar+1)],
  start=c(1973,01,01),end=c(2021,11,01),
  frequency=12)
```

```{r}
#Plot ACF and PACF for industrial energy consumption & industrial energy loss
par(mfrow=c(1,4))
ACF_I_consumption <- Acf(I_electricity_consumption$Industrial.Primary.Energy, lag = 40, plot = TRUE,main="ACF of Industrial Sector Consumption")
PACF_I_consumption <- Pacf(I_electricity_consumption$Industrial.Primary.Energy, lag = 40, plot = TRUE,main="PACF of Industrial Sector Consumption")
ACF_I_loss <- Acf(I_electricity_consumption$Industrial.Energy.Losses, lag = 40, plot = TRUE,main="ACF of Industrial Sector Losses")
PACF_I_loss <- Pacf(I_electricity_consumption$Industrial.Energy.Losses, lag = 40, plot = TRUE,main="PACF of Industrial Sector Losses")
```
