---
title: "TSA_FinalProject_Coding"
output: html_document
date: "2023-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## install packages

```{r package, message=FALSE, include=FALSE}
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(kableExtra)
library(knitr)
library(forecast)

```

## Dataset Information

```{r}
#Import the excel file
library(readxl)
electricity_consumption <- read_excel(path="../Table_2.1a__Energy_Consumption__Residential_-_Commercial_-_and_Industrial_Sectors.xlsx",skip = 10, sheet="Monthly Data",col_names=TRUE) 

#delete the first row with units
electricity_consumption<-electricity_consumption[-1,]

#Keep two columns: primary energy consumption and energy losses for three sectors, respectively
electricity_consumption<-electricity_consumption[,-3][,-3][,-4]
electricity_consumption<-electricity_consumption[,-5][,-5][,-6]
electricity_consumption<-electricity_consumption[,-7][,-7][,-8]

#Round to three digits for all columns
class(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)
electricity_consumption$`Primary Energy Consumed by the Residential Sector` <- as.numeric(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)
round(electricity_consumption$`Primary Energy Consumed by the Residential Sector`,digits=3)

class(electricity_consumption$`Residential Sector Electrical System Energy Losses`)
electricity_consumption$`Residential Sector Electrical System Energy Losses` <- as.numeric(electricity_consumption$`Residential Sector Electrical System Energy Losses`)
round(electricity_consumption$`Residential Sector Electrical System Energy Losses`,digits=3)

class(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)
electricity_consumption$`Primary Energy Consumed by the Commercial Sector` <- as.numeric(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)
round(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`,digits=3)

class(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)
electricity_consumption$`Commercial Sector Electrical System Energy Losses` <- as.numeric(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)
round(electricity_consumption$`Commercial Sector Electrical System Energy Losses`,digits=3)

class(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)
electricity_consumption$`Primary Energy Consumed by the Industrial Sector` <- as.numeric(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)
round(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`,digits=3)

class(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
electricity_consumption$`Industrial Sector Electrical System Energy Losses` <- as.numeric(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
round(electricity_consumption$`Industrial Sector Electrical System Energy Losses`,digits=3)

colnames(electricity_consumption)<-c('Date','Residential.Primary.Energy','Residential.Energy.Losses','Commercial.Primary.Energy','Commercial.Energy.Losses','Industrial.Primary.Energy','Industrial.Energy.Losses')

head(electricity_consumption)
```

```{r}
#Transform to time series data
Date_frame<-as.Date(electricity_consumption$Month)
ts_electricity <- ts(electricity_consumption[,2:7],start=c(1973,01),end=c(2022,11),frequency=12)

#Summary of time series data
summary_data<-summary(ts_electricity)
summary_data

#Calculate means, standard deviations, and ranges for each variable

mean_R1<-mean(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)
sd_R1<-sd(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)
range_R1<-max(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)-min(electricity_consumption$`Primary Energy Consumed by the Residential Sector`)

mean_R2<-mean(electricity_consumption$`Residential Sector Electrical System Energy Losses`)
sd_R2<-sd(electricity_consumption$`Residential Sector Electrical System Energy Losses`)
range_R2<-max(electricity_consumption$`Residential Sector Electrical System Energy Losses`)-min(electricity_consumption$`Residential Sector Electrical System Energy Losses`)

mean_C1<-mean(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)
sd_C1<-sd(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)
range_C1<-max(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)-min(electricity_consumption$`Primary Energy Consumed by the Commercial Sector`)

mean_C2<-mean(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)
sd_C2<-sd(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)
range_C2<-max(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)-min(electricity_consumption$`Commercial Sector Electrical System Energy Losses`)

mean_I1<-mean(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)
sd_I1<-sd(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)
range_I1<-max(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)-min(electricity_consumption$`Primary Energy Consumed by the Industrial Sector`)

mean_I2<-mean(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
sd_I2<-sd(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)
range_I2<-max(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)-min(electricity_consumption$`Industrial Sector Electrical System Energy Losses`)

# Create a table
summary_df <- data.frame(
  Variable = c("Residential_Sector_Consumption", "Residential_Sector_Losses", "Commercial_Sector_Consumption", "Commercial_Sector_Losses", "Industrial_Sector_Consumption", "Industrial_Sector_Losses"),
  Mean = c(584.60,662.45,345.14,605.47,1806.54,576.51),
  StandardDeviation = c(321.54,185.03,146.90,178.14,146.71,68.86),
  Range = c(1295.71,819.46,538.31,723.97,800.93,371.76)
)
  
summary_table<-kbl(summary_df, caption="Summary Statistics Table used with ts() function") %>% 
  kable_styling(full_width = FALSE, position = "center",latex_options = "hold_position")

summary_table
```


## Initial Plots

```{r}
#plot for electricity consumption
cols<-c("residential"="orange","commercial"="steelblue","industrial"="yellow")
ts_plot<-ggplot(electricity_consumption)+
  geom_line(aes(x=Date,y=Residential.Primary.Energy,color="residential"))+
  ylab(paste0("Electricity Consumption (Trillion BTU)",sep="")) +
  xlab("Date")+
  geom_line(aes(x=Date,y=Commercial.Primary.Energy,color="commercial"))+
  geom_line(aes(x=Date,y=Industrial.Primary.Energy,color="industrial"))+
  labs(title="Primary Electricity Consumption of Three Sectors")+
  scale_color_manual(name="series",values = cols)
plot(ts_plot)
          
#plot for energy losses
cols1<-c("residential"="green","commercial"="red","industrial"="blue")
ts_plot1<-ggplot(electricity_consumption)+
  geom_line(aes(x=Date,y=Residential.Energy.Losses,color="residential"))+
  ylab(paste0("Energy Losses (Trillion BTU)",sep="")) +
  xlab("Date")+
  geom_line(aes(x=Date,y=Commercial.Energy.Losses,color="commercial"))+
  geom_line(aes(x=Date,y=Industrial.Energy.Losses,color="industrial"))+
  labs(title="Energy Losses of Three Sectors")+
  scale_color_manual(name="series",values = cols1)
plot(ts_plot1)

```

###Residential Sector
```{r}
#Create a dataframe only for Residential
electricity_consumption_R<-electricity_consumption[,-4:-7]
colnames(electricity_consumption_R)<-c('Date','Primary.Energy','Energy.Losses')

#Inspect data
#head(electricity_price)
nvar <- ncol(electricity_consumption_R) - 1
nobs <- nrow(electricity_consumption_R)

#transform to time series data
ts_electricity_R <- ts(
  electricity_consumption_R[,2:(nvar+1)],
  start=c(1973,01,10),end=c(2022,11,01),
  frequency=12)

```

```{r}
# Create a variable with number of steps ahead we will forecast
n_for <- 12 #play with this, change to 12, 23, 2 and check of the best model is still the same

ts_forPE_R <- ts(ts_electricity_R[1:(nobs-n_for),"Primary.Energy"],
               start=c(1973,01,10),
               frequency=12)
last_PE_obs_R <- ts_electricity_R[(nobs-n_for+1):nobs,"Primary.Energy"] 

ts_forEL_R <- ts(ts_electricity_R[1:(nobs-n_for),"Energy.Losses"],
               start=c(1973,01,10),
               frequency=12)
last_EL_obs_R <- ts_electricity_R[(nobs-n_for+1):nobs,"Energy.Losses"] 
```


##ACF and PACF Plots
```{r}
#ACF and PACF plots
par(mfrow=c(1,4))
ACF_Plot <- Acf(electricity_consumption_R$Primary.Energy, lag = 40, plot = TRUE,main="ACF of Residential Sector Consumption")
PACF_Plot <- Pacf(electricity_consumption_R$Primary.Energy, lag = 40, plot = TRUE,main="PACF of Residential Sector Consumption")
ACF_Plot <- Acf(electricity_consumption_R$Energy.Losses, lag = 40, plot = TRUE,main="ACF of Residential Sector Losses")
PACF_Plot <- Pacf(electricity_consumption_R$Energy.Losses, lag = 40, plot = TRUE,main="PACF of Residential Sector Losses")
```

## Model 1: Arithmetic mean
```{r}
#mean model for primary energy consumption
MEAN_seas1 <- meanf(y = ts_forPE_R, h = n_for)  
checkresiduals(MEAN_seas1)
plot(MEAN_seas1)

#mean model for energy losses
MEAN_seas2 <- meanf(y = ts_forEL_R, h = n_for)  
checkresiduals(MEAN_seas2)
plot(MEAN_seas2)
```

## Model 2: Seasonal naive
```{r}
#seasonal naive model for primary energy consumption
SNAIVE_seas1 <- snaive(ts_forPE_R, h=n_for, holdout=FALSE)
checkresiduals(SNAIVE_seas1)
plot(SNAIVE_seas1)

#seasonal naive model for energy losses
SNAIVE_seas2 <- snaive(ts_forEL_R, h=n_for, holdout=FALSE)
checkresiduals(SNAIVE_seas2)
plot(SNAIVE_seas2)
```

## Model 3: SARIMA
```{r}
##SARIMA model for primary energy consumption
SARIMA_autofit1 <- auto.arima(ts_forPE_R)
checkresiduals(SARIMA_autofit1)

SARIMA_for1 <- forecast(SARIMA_autofit1,h=n_for)
plot(SARIMA_for1)

##SARIMA model for energy losses
SARIMA_autofit2 <- auto.arima(ts_forEL_R)
checkresiduals(SARIMA_autofit2)

SARIMA_for2 <- forecast(SARIMA_autofit2,h=n_for)
plot(SARIMA_for2)
```

## Fitting State Space Models to the original (seasonal) series
## Model 4: SS Exponential smoothing

```{r}
##SSEAS model for primary energy consumption
SSES_seas1 <- es(ts_forPE_R,model="ZZZ",h=n_for,holdout=FALSE)
plot(SSES_seas1)
checkresiduals(SSES_seas1)

##SSEAS model for energy losses
SSES_seas2 <- es(ts_forEL_R,model="XXX",h=n_for,holdout=FALSE)
plot(SSES_seas2)
checkresiduals(SSES_seas2)
```

### Model 5: SS with StructTS()

Now let's try to fit one of the three SS models we learned: local level, local linear trend and the basic structure model (BSM). Since we have seasonality we need to go with the BSM. We will use the `StructTS()` function from package `forecast`.
The function takes a few arguments, here are the important ones:

**x**  time series object
**type** type of the model. If `type ='level'` R will fit a local level model. If `type ='trend'` R will fit a local linear trend model and if `type ='BSM'` R will fit a basic structural model. If argument is omitted, a BSM is used for a time series with frequency > 1, and a local trend model otherwise.
**fixed** optional numeric vector with values for the components variances in the following order $fixed=c(\sigma^2_\eta,\sigma^2_\xi,\sigma^2_\omega,\sigma^2_\epsilon)$. The first variance is with respect to level, second is trend, third is seasonal and fourth is observation. If supplied, only NA entries in fixed will be varied. This argument is useful for setting variances to zero.


```{r}
#For primary energy consumption
SS_seas1 <- StructTS(ts_forPE_R,
                    type="BSM",fixed=c(NA,NA,NA,NA))
checkresiduals(SS_seas1)

#For energy losses
SS_seas2 <- StructTS(ts_forEL_R,
                    type="BSM",fixed=c(NA,NA,NA,NA))
checkresiduals(SS_seas2)

#Generating forecasts
SS_for1 <- forecast(SS_seas1,h=n_for)
plot(SS_for1)

SS_for2 <- forecast(SS_seas2,h=n_for)
plot(SS_for2)
```


## Checking accuracy of the five models

```{r}
#Model 1: Arithmetic mean
MEAN_scores1 <- accuracy(MEAN_seas1$mean,last_PE_obs_R)
MEAN_scores2 <- accuracy(MEAN_seas2$mean,last_EL_obs_R)

#Model 2: Seasonal naive 
SNAIVE_scores1 <- accuracy(SNAIVE_seas1$mean,last_PE_obs_R)
SNAIVE_scores2 <- accuracy(SNAIVE_seas2$mean,last_EL_obs_R)

# Model 3:  SARIMA 
SARIMA_scores1 <- accuracy(SARIMA_for1$mean,last_PE_obs_R)
SARIMA_scores2 <- accuracy(SARIMA_for2$mean,last_EL_obs_R)

# Model 4:  SSES
SSES_scores1 <- accuracy(SSES_seas1$forecast,last_PE_obs_R)
SSES_scores2 <- accuracy(SSES_seas2$forecast,last_EL_obs_R)

# Model 5:  BSM 
SS_scores1 <- accuracy(SS_for1$mean,last_PE_obs_R)
SS_scores2 <- accuracy(SS_for2$mean,last_EL_obs_R)
```

## Compare performance metrics

````{r}
#create data frame for primary energy consumption
seas_scores1 <- as.data.frame(rbind(MEAN_scores1, SNAIVE_scores1, SARIMA_scores1,SSES_scores1,SS_scores1))
row.names(seas_scores1) <- c("MEAN", "SNAIVE","SARIMA","SSES","BSM")

#choose model with lowest RMSE
best_model_index1 <- which.min(seas_scores1[,"RMSE"])
cat("The best model by RMSE is:", row.names(seas_scores1[best_model_index1,]))                       
                            
```

```{r}

kbl_PE_R<-kbl(seas_scores1, 
      caption = "Forecast Accuracy for Residential Primary Energy Consumption's Seasonal Data",
      digits = array(5,ncol(seas_scores1))) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  kable_styling(latex_options="striped", stripe_index = which.min(seas_scores1[,"RMSE"]))

kbl_PE_R
```

````{r}
#create data frame for residential energy losses
seas_scores2 <- as.data.frame(rbind(MEAN_scores2, SNAIVE_scores2, SARIMA_scores2,SSES_scores2,SS_scores2))
row.names(seas_scores2) <- c("MEAN", "SNAIVE","SARIMA","SSES","BSM")

#choose model with lowest RMSE
best_model_index2 <- which.min(seas_scores2[,"RMSE"])
cat("The best model by RMSE is:", row.names(seas_scores2[best_model_index2,]))                       
                            
```

```{r}

kbl_EL_R<-kbl(seas_scores2, 
      caption = "Forecast Accuracy for Residential Enegy Losses' Seasonal Data",
      digits = array(5,ncol(seas_scores2))) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  kable_styling(latex_options="striped", stripe_index = which.min(seas_scores2[,"RMSE"]))

kbl_EL_R
```

## Plotting everything together

```{r}
#Residential Primary Energy
autoplot(ts_electricity_R[,"Primary.Energy"]) +
  autolayer(MEAN_seas1, PI=FALSE, series="Mean") +
  autolayer(SNAIVE_seas1, PI=FALSE, series="Naïve") +
  autolayer(SARIMA_for1,PI=FALSE, series="SARIMA") +
  autolayer(SSES_seas1$forecast, series="SSES") +
  autolayer(SS_for1,PI=FALSE,series="BSM") + 
  xlab("Month") + ylab("Electricity Consumption (Trillion BTU)") +
  guides(colour=guide_legend(title="Forecast"))
```

```{r}
#Residential Energy Losses
autoplot(ts_electricity_R[,"Energy.Losses"]) +
  autolayer(MEAN_seas2, PI=FALSE, series="Mean") +
  autolayer(SNAIVE_seas2, PI=FALSE, series="Naïve") +
  autolayer(SARIMA_for2,PI=FALSE, series="SARIMA") +
  autolayer(SSES_seas2$forecast, series="SSES") +
  autolayer(SS_for2,PI=FALSE,series="BSM") + 
  xlab("Month") + ylab("Electricity Losses (Trillion BTU)") +
  guides(colour=guide_legend(title="Forecast"))
```

